---
title: "Statistical Inference Course Project - Part One"
author: "Fabio Barros"
date: "January, 2016"
output: pdf_document
---

# Overview


# Variables 
First of all, let's set the variables used in this simulation.
```{r}
# Number of simulations
n.sims <- 2000
# Rate parameter
lambda <- 0.2
# Vector of samples
exp.sample <- NULL
# Size of samples
n.sample <- 40
# Vector of x bar values
x.bar <- NULL
```

# Theoretical values
Let's calculate the theoretical values
```{r}
# Calculates the mean of x bar values 
mean.x.bar <- 1 / lambda
# Standard deviation of x bar values
s.x.bar <- (1 / lambda) * (1 / sqrt(n.sample))
# Variance of x bar values
var.x.bar <- s.x.bar ^ 2
```
$E[\overline{X}] = \frac {1}{\lambda} = \frac {1}{0.2} = `r mean.x.bar`$

$s = \frac{\sigma}{\sqrt{n}} = \frac {1}{\lambda} \times \frac {1}{\sqrt{n}} = \frac {1}{0.2} \times \frac {1}{\sqrt{40}} = `r s.x.bar`$

$Var[\overline{X}] = s^ 2 = `r s.x.bar` ^ 2 = `r var.x.bar`$

# Simulations
```{r}
# Sets seed
set.seed(12)
# Generates 20K simulations of 40 samples
for (i in 1:n.sims) {
  # Generates the exponential sample 
  exp.sample <- rexp(n = n.sample, rate = lambda)
  # Creates the distribution of x bar values
  x.bar <- c(x.bar , mean(exp.sample))
}
```

# Sample Mean versus Theoretical Mean

```{r}
# Calculate the cumulative means
acc.mean <- cumsum(x.bar)
for (i in 1:n.sims) {acc.mean[i] <- acc.mean[i] / i}   
# Plot the cumulative means
plot(acc.mean, type = "l", col = "blue", xlab = "Simulations", ylab = "Mean of sample means")
abline(h = mean.x.bar, col = "red", lwd = 2, lty = 2)
```

# Sample Variance versus Theoretical Variance

```{r}
# Calculate the cumulative variances
acc.var <- NULL
for (i in 1:n.sims) {acc.var[i] <- var(x.bar[1:i])}
# Plot the cumulative variances
plot(acc.var, type = "l", col = "darkgreen", xlab = "Simulations", ylab = "Variance of sample means")
abline(h = var.x.bar, col = "red", lwd = 2, lty = 2)
```

# Distribution
```{r}
# Creates a sample of normal distribution values
norm.dist <- rnorm(n = n.sims, mean = mean.x.bar, sd = s.x.bar)
# Plots histogram
hist(x.bar, breaks = 50, probability = TRUE, col = "lightblue", xlab = "",
     main = "Distribution of averages of samples with lambda = 0.2")
lines(density(x.bar), col = "blue", lwd = 2)
abline(v = mean(x.bar), col = "red", lwd = 2)
lines(density(norm.dist), col = "red", lwd = 2, lty = 2)
```